Measuring Hash Trustworthiness via Collision Utility Metrics: Logical Cryptanalysis of MD4

### 2017-12-05 dsn-2018/bare_conf.tex
DSN changes

\section{Intuition on the Utility of a Collision} \label{Sec:Intuition}

\textbf{TODO - audit introductory paragraphs}

We seek to build new intuition about collisions in hash functions
and particularly collisions in MD4. By developing techniques using logical
cryptanalysis, we seek to complement and extend the results from
differential cryptanalysis. In particular, we wish to use logical cryptanalysis
to discuss the impacts of collisions in real-world systems. In this paper,
we introduce and give justification for general techniques, and present some
results of these techniques.

To begin, real world systems are most vulnerable to two attacks: the preimage
and second preimage. For systems relying on password validation via
comparing two hashes, preimage attacks would allow inverting the hash,
increasing the risk associated with losing a database of hashed passwords.
On the other hand, systems relying on message validation, such as signature
checks or file integrity checks would be easily broken with fast second
preimage techniques. Both of these attacks are intractable, thus protecting
real-world systems. However, much research has been devoted to collision
attacks, producing efficient results.

Thus, we seek to evaluate the relative difference between a collision
attack and second preimage attacks, to give some measure of distance between
them. Several high profile attacks against hash function usage have been
made. For MD5, these include cloning a CA certificate \cite{Stevens2009}
and creating malicious executables \cite{MD5Executables}. For SHA-1, a pair of
PDFs with the same hash \cite{cryptoeprint:2017:190}. However, many of these
techniques exploit not the hash, but the flexibility of the file format,
relying instead on embedding arbitrary, collidable data and later checking for
the presence of one of the matching blocks. This is especially obvious in the
MD5 executable and SHA-1 PDF attacks. The work by M. Stevens to perform the
CA Collision, on the other hand, was a chosen prefix attack extending
previous work from 2007 \cite{Stevens2007}.

Seeing the need to evaluate how much flexibility exists in a collision,
we propose the following metrics for evaluating the utility of a collision
class:
\begin{enumerate}
    \item The number of unique differentials a collision class has.
    \item The number of unit-step neighbors a collision class has.
    \item The maximum count of zeros in a the binary representation of a
        colliding block (and likewise with ones).
    \item Whether there exists a block which collides under multiple initial
        values.
    \item Whether or not zero, one, or both of the blocks in a collision may
        be of ASCII values under any input block difference.
\end{enumerate}
Note that the first three are quantitative measures providing some measure of
flexibility of a collision class, whereas the latter two are merely looking
for a single witness having such a property. Depending on the scenario,
specific properties of a collision may be of more interest than others.

The first metric evaluates the flexibility of the differential path. A
differential path with more flexibility will have more differentials which
produce blocks with the given differential path. More differentials implies
a greater flexibility in choice of colliding block, and possibly allows
for multiple collisions for a given colliding block. Furthermore, a collision
with more differentials is more likely to satisfy the last metric, having a
pair of blocks--both ASCII--which produce a collision.

@@

The fourth metric evaluates the utility of a collision when the internal state
of the hash function is unknown. If a collision occurs under multiple initial
values, this could be used to attack some systems where user provided input is
appended to unknown data and then hashed. If a block collides under a suitably
large number of initial values, the attack becomes highly likely to occur
successfully. However, measuring the exact number of initial values a block
collides under is beyond the scope of this work.

The last metric is similar to the third in that it looks for specific bit
patterns in a collision. ASCII is one example of a widely used constraint
system. Further examples, such as JSON, XML, etc., may likewise be
supplemented based on the specifics of the system.

>>>>

\section{Flexibility and the Utility of a Collision Class} \label{Sec:Intuition}

We seek to build new intuition about collision classes in hash functions
and particularly collisions in MD4. By developing techniques using logical
cryptanalysis, we complement and extend the results from
differential cryptanalysis. In particular, we use logical cryptanalysis
to discuss the impacts of collisions in real-world systems, and to
give justification for general techniques, and the
results of these techniques.

Real world systems are most vulnerable to two attacks: preimage
and second preimage. For systems relying on password validation via
comparing two hashes, preimage attacks would allow inverting the hash,
increasing the risk associated with losing a database of hashed passwords.
On the other hand, systems relying on message validation, such as signature
checks or file integrity checks would be easily broken with fast second
preimage techniques. Both of these attacks are theoretically intractable, thus protecting
real-world systems. In practice, however, research efforts have been
successfully devoted to the production of collision attacks, producing
efficient results.  These attacks have real consequences for
trustworthy systems, and evaluating their impact and risk is an
important and underserved area of the existing literature.
We seek to evaluate trustworthiness by building a measure of the relative difference between a collision
attack and second preimage attacks. The literature describes a number
of successful high profile attacks against popular, and industrially
relevant hash functions. 
For MD5, these include cloning a CA certificate \cite{Stevens2009}
and creating malicious executables \cite{MD5Executables}. For SHA-1, a pair of
PDFs with the same hash was successfully produced \cite{cryptoeprint:2017:190}. Many of these
techniques, however, exploit not the hash but the flexibility of the file format
relying instead on embedding arbitrary, collidable data and later checking for
the presence of one of the matching blocks. This is especially obvious in the
MD5 executable and SHA-1 PDF attacks. The work by M. Stevens to perform the
CA Collision, on the other hand, was a chosen prefix attack extending
previous work from 2007 \cite{Stevens2007}.

Evaluating flexibility becomes important as it serves as a
contrapositive to trustworthiness.  A highly flexible attack implies a
larger set of families of collision classes, and decreasing the trust
that should be placed in a hash function. To evaluate how much flexibility exists in a collision,
we propose the following metrics for evaluating the utility of a collision
class:
\begin{enumerate}
    \item The number of unique differentials a collision class has.
    \item The number of unit-step neighbors a collision class has.
    \item The maximum count of zeros in a the binary representation of a
        colliding block (and likewise with ones).
    \item Whether there exists a block which collides under multiple initial
        values.
    \item Whether or not zero, one, or both of the blocks in a collision may
        be of ASCII values under any input block difference.
\end{enumerate}
Note that the first three are quantitative measures providing some measure of
flexibility of a collision class, whereas the latter two are merely looking
for a single witness having such a property. Depending on the scenario,
specific properties of a collision may be of more interest than others.

The first metric evaluates the flexibility of the differential path. A
differential path with more flexibility will have more input-block differentials which
produce blocks with the given differential path. More differentials implies
a greater flexibility in choice of colliding block, and possibly allows
for multiple collisions for a given colliding block. Furthermore, a collision
with more differentials is more likely to satisfy the last metric, having a
pair of blocks --- both ASCII --- which produce a collision.

@@

The fourth metric evaluates the utility of a collision when the internal state
of the hash function is unknown. If a collision occurs under multiple initial
values, this could be used to attack some systems where the input
provided by a user is
appended to unknown data and then hashed. If a block collides under a suitably
large number of initial values, the attack becomes highly likely to occur
successfully. Measuring the exact number of initial values a block
collides under, however, is beyond the scope of this work.

The final metric is similar to the third in that it looks for specific bit
patterns in a collision. ASCII is one example of a widely used constraint
system. Further examples, such as JSON, XML, etc., may likewise be
supplemented based on the specifics of the system, and the domain of
its application and implementation.
